---
title: "Factor Analysis"
author: "Neehaar Gandhi, David Liu, Josh Brooks"
date: "4/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(psych)
library(rela)
library(corrplot)
library(PerformanceAnalytics)
```

```{r}
#read data
PD <- read.csv("data.csv", header=TRUE, stringsAsFactors = FALSE)

#isolate only complete cases
PD <- PD[complete.cases(PD),]

#choose a subset of the data that makes sense  

#data <- PD[,c(4,6,9,16,17,18,19,29,30,31)]
#data[,5] <- as.numeric(data[,5])
#the above was our first attempt to choose good indicators. the second attempt, which is below, seemed more successful. 

data <- PD[,c(29,30,31,32,33,37,38,39,40,47,49,55,56,57)]
data
```

1. Look through indicators (questions). Think about which indicators might be related through latent factors.
*We chose the following skill 14 rating variables as indicators. Among this list are a number of traits that we suspect to be related to offensive position success, as well as a number of traits which we suspect to be related to defensive position success.*  

2. Compute the correlation matrix between all indicators (you may want to do this in batches). Comment on relationships you do/do not observe.
```{r}
data1 <- data[,1:length(data)]
cor(data1)
corrplot(cor(data1),method = "ellipse", order="hclust", tl.col = "black", tl.cex=.8)
chart.Correlation(data1, histogram=TRUE, pch=19)
```
*Our correlation matrix and its visual aids show that we have a number of predictors that are correlated in similar ways, suggesting the viability of finding some underlying factors affecting variation. Some correlation magnitudes are smaller, but many are somewhat large, between 0.6 and 0.9.*  

3. Compute KMO or other measure (i.e. just look at matrix produced above) to comment on suitability of data for factor analysis.
```{r}
library(rela)
datamat <- data.matrix(data1)
#paf(datamat, eigcrit=1, convcrit=.001)
output1 <- paf(datamat, eigcrit=1, convcrit=.001)
print(output1$KMO)
summary(output1)
```
*The selected indicators give us a Kaiser-Meyer-Olkin output of 0.83571, which is well above the acceptable range, and verging on good -- "meritorious", according to Kaiser's rating scale. Generally speaking, this means that there is a relatively large degree of common variance among our variables, and thus our data are suitable enough for facter analysis.*  

*We initially tried a different and less judiciously chosen set of predictors, which gave a KMO value of about 0.65. This number, while also probably acceptable for factora analysis, is on the low side. As such, we decided to choose a better sample.*  

4. Use Principle Components (or appropriate option in Factor Analysis) to decide on a number of latent factors. You can use Scree Plot, eigenvalue>1, or parallel analysis. 

```{r}
#principal components analysis
#chose to do PCA on the covariance matrix because the variable scales are similar (they are 0-100)
fit <- princomp(data, cor=FALSE)
summary(fit)
loadings(fit)

#eigenvalue > 1
print(summary(fit),digits=2,loadings=fit$loadings,cutoff=0)
#all 13 components have a stdev and variance (eigenvalue) of greater than one, perhaps use a different metric to determine number of latent factors. 

#scree plot
plot(fit, type="lines", main="Scree Plot of FIFA Data",col="red",lwd=2,pch=19,cex=1.2)
#seems to be an elbow at the 5th component
```

5. Perform a series of factor analysies using orthogonal models. First, try at least two extration methods (choose from Principle Components, Principle Axis Factoring, Iterative Principle Components, Maximum Likelihood), Use some method for comparing extraction methods to choose a 'best' method (i.e. RMSR or # residuals greater than .05).

#### Perform Factor Analysis using iterative PCA with Varimax Rotation

```{r}
fact3 <- fa(data, nfactors=3, rotate="varimax", SMC=FALSE, fm="pa")
fact3
```

Loading plot for first two factors.

```{r}
plot(fact3$loadings, pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact3$loadings, labels=names(data),cex=0.8)
```

Correlation Matix Analyses

```{r}
#get reproduced correlation matrix
repro3 <- fact3$loadings%*%t(fact3$loadings)
#residual correlation matrix
resid3 <- cor(data)-repro3
round(resid3,2)

#get root-mean squared residuals - again, in output above
len <- length(resid3[upper.tri(resid3)])
RMSR3 <- sqrt(sum(resid3[upper.tri(resid3)]^2)/len)
RMSR3

#get proportion of residuals greater than 0.05 in absolute value
sum(rep(1,len)[abs(resid3[upper.tri(resid3)])>0.05])/len

```

#### Factor Analysis using PAF with Varimax Rotation

```{r}
#this uses the fa() function in the psych package.  Note that this fails with only 2 factors
fact2 <- fa(data, nfactors=3, rotate="varimax", fm="pa")
fact2
```


Loading plot for first two factors

```{r}
plot(fact2$loadings, pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact2$loadings, labels=names(data),cex=0.8)
```


Correlation matrix analysis

```{r}
#get reproduced correlation matrix
repro2 <- fact2$loadings%*%t(fact2$loadings)
#residual correlation matrix
resid2 <- cor(data)-repro2
round(resid2,2)

#get root-mean squared residuals - already provided in output actually
len <- length(resid2[upper.tri(resid2)])
RMSR2 <- sqrt(sum(resid2[upper.tri(resid2)]^2)/len)
RMSR2

#get proportion of residuals greater than 0.05 in absolute value
sum(rep(1,len)[abs(resid2[upper.tri(resid2)])>0.05])/len

```

6. Once you've chosen an extraction method, try a varimax and/or a quartimax rotation. Pick one of these rotations and discuss the interpretation of the final factors. Make one or more loading plots as appropriate.

```{r}

```