---
title: "Factor Analysis"
author: "Neehaar Gandhi, David Liu, Josh Brooks"
date: "4/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(psych)
library(rela)
```

```{r}
#read data
PD <- read.csv("data.csv", header=TRUE, stringsAsFactors = FALSE)

#isolate only complete cases
PD <- PD[complete.cases(PD),]

#choose a subset of the data that 
data <- PD[,c(4,6,9,16,17,18,19,29,30,31)]

#data fix
data[,5] <- as.numeric(data[,5])
```

1. Look through indicators (questions). Think about which indicators might be related through latent factors.

2. Compute the correlation matrix between all indicators (you may want to do this in batches). Comment on relationships you do/do not observe.
```{r}
data1 <- data[,1:length(data)]
cor(data1)
```

3. Compute KMO or other measure (i.e. just look at matrix produced above) to comment on suitability of data for factor analysis.
```{r}
library(rela)
datamat <- data.matrix(data1)
paf(datamat, eigcrit=1, convcrit=.001)
```

4. Use Principle Components (or appropriate option in Factor Analysis) to decide on a number of latent factors. You can use Scree Plot, eigenvalue>1, or parallel analysis. 

```{r}
fact3 <- fa(data, nfactors=3, rotate="varimax", SMC=FALSE, fm="pa")
fact3
```


Loading plot for first two factors.

```{r}
plot(fact3$loadings, pch=18, col='red')
abline(h=0)
abline(v=0)
text(fact3$loadings, labels=names(data),cex=0.8)
```


Correlation Matix Analyses

```{r}
#get reproduced correlation matrix
repro3 <- fact3$loadings%*%t(fact3$loadings)
#residual correlation matrix
resid3 <- cor(data)-repro3
round(resid3,2)

#get root-mean squared residuals - again, in output above
len <- length(resid3[upper.tri(resid3)])
RMSR3 <- sqrt(sum(resid3[upper.tri(resid3)]^2)/len)
RMSR3

#get proportion of residuals greater than 0.05 in absolute value
sum(rep(1,len)[abs(resid3[upper.tri(resid3)])>0.05])/len

```

5. Perform a series of factor analysies using orthogonal models. First, try at least two extration methods (choose from Principle Components, Principle Axis Factoring, Iterative Principle Components, Maximum Likelihood), Use some method for comparing extraction methods to choose a 'best' method (i.e. RMSR or # residuals greater than .05).

6. Once you've chosen an extraction method, try a varimax and/or a quartimax rotation. Pick one of these rotations and discuss the interpretation of the final factors. Make one or more loading plots as appropriate.