---
title: "SDS 363 - Final Project"
author: "Neehaar Gandhi, David Liu, Josh Brooks"
date: "4/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction
This project analyzes the data that drives the 2018 edition of the officially licensed Fédération Internationale de Football Association (International Federation of Association Football) video game, titled FIFA 18, produced by Electronic Arts (EA). Association football (soccer) is by far the most popular sport in the world, so it should come as no surprise that the game has became something of a cultural icon. To date, the FIFA franchise is the sixth-best selling video game series in history, with its popularity likely to continue rising. A new version of the game is released once per year, making the 2018 edition of the game the twenty-fifth installment. 

Within the electronic entertainment industry, sports video games present a unique, data-driven challenge. Gamers want to be titillated and drawn in, while rapid sports fans simply will never tolerate misrepresentation of their favorite players. This tension -- between manufacturing a maximally effective fictitious world and satisfying players’ wish to play with real people -- leads to a wealth of data-driven challenges. This is the central consideration that underscores our analysis. We will use the game’s player rating data, which is described in detail in the following section, to answer questions about the relationships between soccer player attributes.

#Design and Primary Questions
One area of interest for us is understanding better how EA producers ultimately come to their “Overall” player ratings, which is often markedly higher than the average of their other skill categories. This leads us to believe that there is an element of subjectivity in determining overall player scores; EA sports determines that some players have the x-factor needed to boost their overall scores, while others do not. Thus, a principal question for us is: *What sets of variables correspond with higher overall scores, and can we use those variables to tell a story about what an ‘x-factor’ might be?* To answer this we will use factor analysis. In so doing we hope we will be able to find a latent factor that might tell a story about what makes players valuable on the field. 

Further, we seek to look for patterns of players within particular position groups. *Within broad positional categories -- back, midfielder, forward -- can we see discernable subgroups of players?* Again, we would certainly expect to see this, given subjective understandings of soccer tactics. Using k-means as a clustering technique, we hope to create a readable picture of positional subgroups.

Finally, we want to look at the interactions between FIFA players’ nationalities and their skills in various areas. In the international soccer community, it is very common to talk about geographic regions as very different. For instance, it has become accepted wisdom to say that Latin American teams play a more skillful, less team-oriented style, while European teams play with more precision and less individual skill. We hope to see whether the data bear this out. In this vein, our third question for analysis will be: *Do different country groups create players that are differently abled in particular facets of the game?* To test this we will use MANOVA and attempt to establish differences in ability means between various nationality groups across a few levels of player ability. 

#Data
```{r, include=FALSE}
#include relevant libraries for project
library(aplpack)
library(fpc)
library(cluster)
library(ape)
library(ggplot2)
library(DBItest)
library(tidyverse)  
library(factoextra)
library(psych)
library(rela)
library(corrplot)
library(PerformanceAnalytics)
library(gridExtra)
library(NbClust)

PD <- read.csv("data.csv", header=TRUE, stringsAsFactors = FALSE)
#isolate only complete cases
PD <- PD[complete.cases(PD),]

#make sure data is numeric where it ought to be numeric (in the original dataset, for some reason some player heights/weights/etc [variables that need to be put in numeric terms], were read as characters [perhaps due to spaces or apostrophes])
PD[,4] <- as.numeric(PD[,4])
PD[,6] <- as.numeric(PD[,6])
PD[,7] <- as.numeric(PD[,7])
PD[,9] <- as.numeric(PD[,9])
PD[,17] <- as.numeric(PD[,17])
PD[,18] <- as.numeric(PD[,18])
```
The dataset is quite large, as a result of the game’s intention of accurately representing every male professional soccer player in the world. While the video game surely requires an even larger ocean of data to operate, we will concern ourselves only with the player rating data -- those data that describe the qualities of individual athletes and their particular traits and skills. 
There are a few broad categories of data in the set. The first type of variable is made up of the information that describes a player and his location. These include physical traits like height and weight, as well as the club for which the player plays professionally, the country for which the player plays in international competition, the player’s wage per game-week, and the like. All of this information is taken directly from publicly available postings by the players’ clubs. Wage is rounded to the nearest thousand, which could change the data somewhat at lower numbers. 

The second, and ultimately the most significant, type of variable in the set comprises ratings for various in-game skills and actions on a 100 point scale. The full list has more than thirty predictors, which will be detailed below. Among them are skilled related to playing offense, playing defense, passing, athleticism, and everyone else that one must do on a soccer field. Some examples are: “Ball Control” (ability to keep possession of the ball and place it in a favorable position for future actions), “Finishing” (generalized fast-twitch ability to accurately kick the ball into the goal after receiving a pass), “Marking” (ability to successfully keep with a player while playing defense, and prevent him from receiving the ball), and “Agility” (a general measure of ability to move quickly over short distances and control one’s body through space). 

These skill ratings are, of course, highly subjective. They do not appear out of thin air; they must be created and compiled by human raters. EA are upfront about their process. They delegate the much of the task of watching and evaluating to a network of over 9000 “data reviewers”, made up of professionals (coaches) and amateurs (qualified fans). The material from the reviewers comes to a group of a few-hundred editors, who compile the data into the rating categories listed below. With this in mind, we can think of these ratings as the results of a large survey, in which soccer fanatics are asked to watch a great deal of game film and give answers to questions about player ability. This is not exactly true -- EA’s producers and editors tweak the data as they see fit, and the sampling of data reviewers is surely not perfectly random -- but thinking of it this way provides a general picture.

The third group of predictors is smaller and somewhere between the two previous. It is made up of skill ratings with only five categorical levels. These include “Skill Moves” and “International Reputation”. We will ignore many these in our analysis -- they are less about soccer, and more about game mechanics. Skill moves is a good example of this category. From 1 to 5, as players go to up higher thresholds of skill rating, they gain the ability to carry out particular in-game actions. However, 

#Descriptive plots, summary statistics
```{r}
summary(PD)
head(PD)

# Frequency Examinations for Categoricals
table(PD$Nationality)
table(PD$Club)

# Plot to give a sense of the variance in number of players produced by different countries
plot(table(PD$Nationality))

#summary histograms for all of the continuous variables using ggplot function
PD[c(4, 6, 7, 9, 11, 12, 13, 16:57)] %>%
  gather(key="key",value="value") %>% 
  ggplot(aes(x=value))+
  geom_density()+
  facet_wrap(vars(key),scales="free")


#boxplots for some of the variables that did not have histograms that resembled an approximately normal distribution
boxplotCols <- c(9, 11, 13, 39, 4, 40, 57, 13)
# boxplotCols <- c(4, 6, 8) # use this to get a sense of plots when u don't want to load all plots
PDNames <- colnames(PD)
for (col in boxplotCols){
  boxplot(PD[, col], main=PDNames[col])
}

# Correlation matrix shown in latter part
```


#Multivariate Analysis and Discussion of Results

###Factor Analysis
```{r}
#subset data that we will use for factor analysis
data <- PD[,c(4, 6, 7, 9, 11, 12, 13, 16:57)]
```

```{r}
#partition into thirds so that correlation plots are more readible
whole <- data[,1:length(data)]
data1 <- data[,1:length(data)/3]
data2 <- data[,(length(data)/3):(2*(length(data)/3))]
data3 <- data[,(2*(length(data)/3)):length(data)]

#compute the correlation matrix between all indicators
corrplot(cor(data1),method = "ellipse", order="hclust", tl.col = "black", tl.cex=.8)
corrplot(cor(data2),method = "ellipse", order="hclust", tl.col = "black", tl.cex=.8)
corrplot(cor(data3),method = "ellipse", order="hclust", tl.col = "black", tl.cex=.8)

# chart.Correlation(data1, histogram=TRUE, pch=19)
# chart.Correlation(data2, histogram=TRUE, pch=19)
# chart.Correlation(data3, histogram=TRUE, pch=19)

```
*We see many highly correlated predictors. There are strong correlations between many sorts of factors, all of which make a great deal of sense when examined individually. Volleying (kicking the ball out of the air and shooting it at the goal) is highly correlated with finishing (general ability to kick the ball into the goal); finishing ability is very negatively correlated with defensive positional ratings (like center back [CB]); tackling ability (the ability to take the ball away from an opposing player) is highly correlated with ability to intercept the ball (to take possesion of a pass intended for an opposing player).*

*Further, our correlation matrices show that we have a number of indicators that are correlated deeply, suggesting the viability of finding some underlying factors affecting variation. Some correlation magnitudes are smaller, but many are somewhat large, between 0.6 and 0.9. For instance, some of the larger correlations have to do with defensive ability such as marking and slide tackling. We also see some high correlations involving indicators that relate to skill.*

*Our interest in multivariate factor analysis of this dataset stems from those relationships which are not necessarily as easily expected. Fast players seem to be rated as worse short passers than long passers; players rated as aggressive are more likely to have high 'composure' ratings; strong tacklers seem to be rated as worse at positioning themselves on the field. Many of these combinations of attributes do not seem to make sense together. We think factor analysis will help us see some of the underlying relationships between traits, and ultimately give us insight into the combinations of skills that make a player 'good', either in the video game or in real life.*

```{r}
#KMO
datamat <- data.matrix(whole)
output1 <- paf(datamat, eigcrit=1, convcrit=.001)
print(output1$KMO)
#summary(output1)
```
*In conducting factor analysis for our data, we first had to determine the suitability by computing KMO (Kaiser-Meyer-Olkin). Further, a test such as Principal Component Analysis (PCA) must be conducted to determine a number of latent factors that we are looking for, and as such, we will compute PCA on the covariance matrix because the variable scales we utilized are similar (ranging from 0-100). The selected indicators give us a Kaiser-Meyer-Olkin output of 0.74211, which is acceptable. We can proceed with Factor Analysis.*

```{r}
#Using PCA to help decide on a number of latent factors
fit <- princomp(whole, cor=FALSE)

#used two different metrics to determine how many latent factors we should use in our analyses

#eigenvalue > 1 to determine number of latent factors
print(summary(fit),digits=2,loadings=fit$loadings,cutoff=0)

#scree plot
plot(fit, type="lines", main="Scree Plot for Factor Analysis of FIFA Data",col="red",lwd=2,pch=19,cex=1.2)
```
*We ran PCA using the covariance matrix because the scales of the variables were identical. By using the eigenvalue>1 measure, we found that components 1-35 had eigenvalues greater than one, and as such determined it best to use the scree plot metric, instead to determine number of latent factors, because this seems to be too many, and not necessarily that helpful.From the scree plot we can see an elbow at 3rd component. As such, we will aim for 3 latent factors in our factor analysis.*

####Below is the actual Factor Analysis Test using Maximum Likelihood with a Varimax Rotation
```{r}
fact_ml <- fa(datamat, nfactors = 3, SMC = FALSE, rotate = "varimax", fm="ml")
fact_ml$PVAL

#Loading plot for first two factors using Maximum Likelihood
plot(fact_ml$loadings, pch=18, col='red', main="ML Loading Plot for First Two Factors",xlab="1st Factor: Ball Control", ylab="2nd Factor: Ball Pressure")
abline(h=0)
abline(v=0)
text(fact_ml$loadings,labels=names(data),cex=0.8)

#Correlation matrix analysis
#get reproduced correlation matrix
repro_ml <- fact_ml$loadings%*%t(fact_ml$loadings)
#residual correlation matrix
resid_ml <- cor(data)-repro_ml
#round(resid_ml,2)
#omitted for the sake of succinctness

#get root-mean squared residuals - already provided in output actually
len <- length(resid_ml[upper.tri(resid_ml)])
RMSR_ml <- sqrt(sum(resid_ml[upper.tri(resid_ml)]^2)/len)
#RMSR_ml
#omitted for the sake of succinctness

#get proportion of residuals greater than 0.05 in absolute value
sum(rep(1,len)[abs(resid_ml[upper.tri(resid_ml)])>0.05])/len
```

####For the sake of comparison, we have also included Factor Analysis using Principal Axis Factoring with a varimax rotation.
```{r}
#this uses the fa() function in the psych package.  Note that this fails with only 2 factors
fact_paf <- fa(datamat, nfactors = 3, rotate = "varimax", SMC = FALSE, fm="pa")
fact_paf$PVAL
fact_paf

#Loading plot for first two factors Principal Axis Factoring
plot(fact_paf$loadings, pch=18, col='red', main="PAF Loading Plot for First Two Factors",xlab="1st Factor: Ball Control", ylab="2nd Factor: Ball Pressure")
abline(h=0)
abline(v=0)
text(fact_paf$loadings, labels=names(data),cex=0.8)

#Correlation matrix analysis
#get reproduced correlation matrix
repro_paf <- fact_paf$loadings%*%t(fact_paf$loadings)
#residual correlation matrix
resid_paf <- cor(data)-repro_paf
#round(resid_paf,2)
#omitted for the sake of succinctness

#get root-mean squared residuals - already provided in output actually
len <- length(resid_paf[upper.tri(resid_paf)])
RMSR_paf <- sqrt(sum(resid_paf[upper.tri(resid_paf)]^2)/len)
#RMSR_paf
#omitted for the sake of succinctness

#get proportion of residuals greater than 0.05 in absolute value
sum(rep(1,len)[abs(resid_paf[upper.tri(resid_paf)])>0.05])/len
```

*Of the two factor analyses methods used, the principal axis factoring extraction method with varimax rotation produced a lower RMSR value (0.14286 vs 0.15986) and also had the smaller proportion of residuals greater than .05. As such, the following analyses will be based off of the principal axis factoring method (the maximum likelihood method is meant only to be used for comparison's sake, and no conclusions will be made from this method).*

*The first latent factor seems to be about handling when in posession of the ball (ball control), and ability to act as an effective attacker. The second latent factor seems to be about ability to act as an effective defender (ball pressure). The third latent factor expresses variation on skills that are general player attributes, such as height, weight, and other characteristics marking athleticism — there are positive values for acceleration, sprint speed and finishing, but negative values for height and weight. Together, these three factors account for 68% of the variance in the data.* 

*A player's overall score seems to be most closely associated with factor one, which we determined to be ability to maintain solid possession of the ball, attack fiercely, and capitalize off of opportunity. Thus, we are able to determine that when EA Sports Analysts are subjectively determining the x-factor, it is very likely that players that are able to attack the ball well and capitalize often will be given a boost when determining their overall rating in the game.*

###K-Means
```{r}
#read data
PD <- read.csv("data.csv", header=TRUE, stringsAsFactors = FALSE)

PD <- PD[complete.cases(PD),]

rowNames <- PD[,"Name"]

PD2 <- PD
rowNames <- PD2[,"Name"]

PD[,c("X", "ID", "Name", "Nationality", "Club", "PreferredFoot", "BodyType", "Position", "InternationalReputatiton", "WeakFoot", "SkillMoves")] <- list(NULL)

PD[,7] <- as.numeric(PD[,7])

#sapply(1:ncol(PD), function(a) {class(PD[,a])})

SPD <- scale(PD)

SPD



#get defensive subroup
PDdef <- PD2[ which(PD2$Position == "CB" | PD2$Position == "RCB" | PD2$Position == "LCB" | PD2$Position == "RB" | PD2$Position == "LB" | PD2$Position == "LCB" | PD2$Position == "RCB"), ]

PDdef[,c("X", "ID", "Name", "Nationality", "Club", "PreferredFoot", "BodyType", "Position", "InternationalReputation", "WeakFoot", "SkillMoves")] <- list(NULL)

PDdef[,6] <- as.numeric(PDdef[,6])

SPDdef <- scale (PDdef)

SPDdef

#get midfield subgroup
PDmid <- PD2[ which(PD2$Position == "CM" | PD2$Position == "CDM" | PD2$Position == "CAM" | PD2$Position == "CDM" | PD2$Position == "CAM" | PD2$Position == "LAM" | PD2$Position == "RAM" | PD2$Position == "LCM" | PD2$Position == "RCM" | PD2$Position == "LM" | PD2$Position == "RM"), ]

PDmid[,c("X", "ID", "Name", "Nationality", "Club", "PreferredFoot", "BodyType", "Position", "InternationalReputation", "WeakFoot", "SkillMoves")] <- list(NULL)

PDmid[,6] <- as.numeric(PDmid[,6])

SPDmid <- scale (PDmid)

SPDmid

#get offensive subgroup
PDoff <- PD2[ which(PD2$Position == "ST" | PD2$Position == "RW" | PD2$Position == "RS" | PD2$Position == "LS" | PD2$Position == "LW" | PD2$Position == "RW" | PD2$Position == "CF" | PD2$Position == "LF" | PD2$Position == "RF"), ]

PDoff[,c("X", "ID", "Name", "Nationality", "Club", "PreferredFoot", "BodyType", "Position", "InternationalReputation", "WeakFoot", "SkillMoves")] <- list(NULL)

PDoff[,6] <- as.numeric(PDoff[,6])

SPDoff <- scale (PDoff)

SPDoff
```

#### run k-means with offensive players
```{r}
fviz_nbclust(SPDoff, kmeans, method = "silhouette") +
labs(title= "Optimal Cluster Number for Offensive Players - Silhouette Method") 
fviz_nbclust(SPDoff, kmeans, method = "wss") +
labs(title= "Optimal Cluster Number for Offensive Players - WSS Method") +
geom_vline(xintercept = 3 , linetype = 2)

nboff <- NbClust(SPDoff, distance = "euclidean", min.nc = 2,
        max.nc = 10, method = "complete", index ="all")

fviz_nbclust(nboff) + theme_minimal()
```

```{r}
offk2 <- kmeans(SPDoff, centers = 2)
offk3 <- kmeans(SPDoff, centers = 3)
offk4 <- kmeans(SPDoff, centers = 4)
offk5 <- kmeans(SPDoff, centers = 5)

offp2 <- fviz_cluster(offk2, geom = "point", data = SPDoff) + 
  ggtitle("k = 2")
offp3 <- fviz_cluster(offk3, geom = "point",  data = SPDoff) + 
  ggtitle("k = 3")
offp4 <- fviz_cluster(offk4, geom = "point",  data = SPDoff) + 
  ggtitle("k = 4")
offp5 <- fviz_cluster(offk5, geom = "point",  data = SPDoff) + 
  ggtitle("k = 5")

marrangeGrob(grobs=list(offp2, offp3, offp4, offp5), ncol=2, nrow = 2)
```

#### run k-means with defensive players
```{r}
fviz_nbclust(SPDdef, kmeans, method = "silhouette") +
labs(title= "Optimal Cluster Number for Defensive Players - Silhouette Method") 
fviz_nbclust(SPDdef, kmeans, method = "wss") +
labs(title= "Optimal Cluster Number for Defensive Players - WSS Method") +
geom_vline(xintercept = 3 , linetype = 2)

nboff <- NbClust(SPDoff, distance = "euclidean", min.nc = 2,
        max.nc = 10, method = "complete", index ="all")

fviz_nbclust(nboff) + theme_minimal()
```

```{r}
defk2 <- kmeans(SPDdef, centers = 2)
defk3 <- kmeans(SPDdef, centers = 3)
defk4 <- kmeans(SPDdef, centers = 4)
defk5 <- kmeans(SPDdef, centers = 5)

defp2 <- fviz_cluster(defk2, geom = "point", data = SPDdef) + 
  ggtitle("k = 2")
defp3 <- fviz_cluster(defk3, geom = "point",  data = SPDdef) + 
  ggtitle("k = 3")
defp4 <- fviz_cluster(defk4, geom = "point",  data = SPDdef) + 
  ggtitle("k = 4")
defp5 <- fviz_cluster(defk5, geom = "point",  data = SPDdef) + 
  ggtitle("k = 5")

marrangeGrob(grobs=list(defp2, defp3, defp4, defp5), ncol=2, nrow = 2)
```

#### run k-means with center midfielder players
```{r}
fviz_nbclust(SPDmid, kmeans, method = "silhouette") +
labs(title= "Optimal Cluster Number for Center Midfield Players - Silhouette Method") 
fviz_nbclust(SPDmid, kmeans, method = "wss") +
labs(title= "Optimal Cluster Number for Center Midfield Players - WSS Method") +
geom_vline(xintercept = 3 , linetype = 2)

nboff <- NbClust(SPDoff, distance = "euclidean", min.nc = 2,
        max.nc = 10, method = "complete", index ="all")

fviz_nbclust(nboff) + theme_minimal()
```


```{r}
midk2 <- kmeans(SPDmid, centers = 2)
midk3 <- kmeans(SPDmid, centers = 3)
midk4 <- kmeans(SPDmid, centers = 4)
midk5 <- kmeans(SPDmid, centers = 5)

midp2 <- fviz_cluster(midk2, geom = "point", data = SPDmid) + 
  ggtitle("k = 2")
midp3 <- fviz_cluster(midk3, geom = "point",  data = SPDmid) + 
  ggtitle("k = 3")
midp4 <- fviz_cluster(midk4, geom = "point",  data = SPDmid) + 
  ggtitle("k = 4")
midp5 <- fviz_cluster(midk5, geom = "point",  data = SPDmid) + 
  ggtitle("k = 5")

marrangeGrob(grobs=list(midp2, midp3, midp4, midp5), ncol=2, nrow = 2)
```

[insert analysis here]

*MANOVA*

#isolate european and latin american subgroups, and make new categorical variable
```{r}
PDeurlat <- PD[ which(PD$Nationality == "Russia" | PD$Nationality == "Germany" | PD$Nationality == "England" | PD$Nationality == "France" | PD$Nationality == "Italy" | PD$Nationality == "Spain" | PD$Nationality == "Ukraine" | PD$Nationality == "Poland" | PD$Nationality == "Romania" | PD$Nationality == "Netherlands" | PD$Nationality == "Argentina" | PD$Nationality == "Bolivia" | PD$Nationality == "Brazil" | PD$Nationality == "Chile" | PD$Nationality == "Colombia" | PD$Nationality == "Ecuador" | PD$Nationality == "Paraguay" | PD$Nationality == "Peru" | PD$Nationality == "Uruguay" | PD$Nationality == "Venezuela"), ]

Region <- ifelse(PDeurlat$Nationality == "Russia" | PDeurlat$Nationality == "Germany" | PDeurlat$Nationality == "England" | PDeurlat$Nationality == "France" | PDeurlat$Nationality == "Italy" | PDeurlat$Nationality == "Spain" | PDeurlat$Nationality == "Ukraine" | PDeurlat$Nationality == "Poland" | PDeurlat$Nationality == "Romania" | PDeurlat$Nationality == "Netherlands","Europe","LatinAmerica")

PDeurlat$Region <- Region

```

#make interaction plots
```{r}

interaction.plot(PDeurlat$Region,PDeurlat$InternationalReputation,PDeurlat$HeadingAccuracy, lwd=3,col=c("red","blue","black","green","purple"),trace.label = "International Reputation",xlab="Region",ylab="Mean of Heading Accuracy",main="Interaction Plot for Heading Accuracy", legend=FALSE)
legend("topleft", c("Poor","Low","Medium","Fair","Good"),bty="n",lty=2,lwd=3,col=c("red","blue","black","green","purple"), title="International Reputation",inset = 0, cex=.8,y.intersp
=.7, horiz = TRUE)

interaction.plot(PDeurlat$Region,PDeurlat$InternationalReputation,PDeurlat$Agility,lwd=3,col=c("red","blue","black","green","purple"),trace.label="International Reputation",xlab="Region",ylab="Mean of Agility",main="Interaction Plot for Agility",legend=FALSE)
legend("topleft", c("Poor","Low","Medium","Fair","Good"),bty="n",lty=2,lwd=3,col=c("red","blue","black","green","purple"), title="International Reputationt",inset = 0, cex=.8,y.intersp
=.7, horiz = TRUE)

interaction.plot(PDeurlat$Region,PDeurlat$InternationalReputation,PDeurlat$Stamina, lwd=3,col=c("red","blue","black","green","purple"),trace.label = "International Reputation",xlab="Region",ylab="Mean of Stamina",main="Interaction Plot for Stamina", legend=FALSE)
legend("topleft", c("Poor","Low","Medium","Fair","Good"),bty="n",lty=2,lwd=3,col=c("red","blue","black","green","purple"), title="International Reputation",inset = 0, cex=.8,y.intersp
=.7, horiz = TRUE)

interaction.plot(PDeurlat$Region,PDeurlat$InternationalReputation,PDeurlat$Jumping, lwd=3,col=c("red","blue","black","green","purple"),trace.label = "International Reputation",xlab="Region",ylab="Mean of Jumping",main="Interaction Plot for Jumping", legend=FALSE)
legend("topleft", c("Poor","Low","Medium","Fair","Good"),bty="n",lty=2,lwd=3,col=c("red","blue","black","green","purple"), title="International Reputation",inset = 0, cex=.8,y.intersp
=.7, horiz = TRUE)

interaction.plot(PDeurlat$Region,PDeurlat$InternationalReputation,PDeurlat$Balance, lwd=3,col=c("red","blue","black","green","purple"),trace.label = "International Reputation",xlab="Region",ylab="Mean of Balance",main="Interaction Plot for Balance", legend=FALSE)
legend("topleft", c("Poor","Low","Medium","Fair","Good"),bty="n",lty=2,lwd=3,col=c("red","blue","black","green","purple"), title="International Reputation",inset = 0, cex=.8,y.intersp
=.7, horiz = TRUE)

interaction.plot(PDeurlat$Region,PDeurlat$InternationalReputation,PDeurlat$SprintSpeed, lwd=3,col=c("red","blue","black","green","purple"),trace.label = "International Reputation",xlab="Region",ylab="Mean of Sprint Speed",main="Interaction Plot for Sprint Speed", legend=FALSE)
legend("topleft", c("Poor","Low","Medium","Fair","Good"),bty="n",lty=2,lwd=3,col=c("red","blue","black","green","purple"), title="International Reputation",inset = 0, cex=.8,y.intersp
=.7, horiz = TRUE)

interaction.plot(PDeurlat$Region,PDeurlat$InternationalReputation,PDeurlat$Acceleration, lwd=3,col=c("red","blue","black","green","purple"),trace.label = "International Reputation",xlab="Region",ylab="Mean of Acceleration",main="Interaction Plot for Acceleration", legend=FALSE)
legend("topleft", c("Poor","Low","Medium","Fair","Good"),bty="n",lty=2,lwd=3,col=c("red","blue","black","green","purple"), title="International Reputation",inset = 0, cex=.8,y.intersp
=.7, horiz = TRUE)
```
*These interaction plots show promise for MANOVA or some sort of generalized linear model.*

#### One-way Manova
```{r}
mod0 <- manova(as.matrix(PDeurlat[,c("Acceleration", "Stamina")]) ~ PDeurlat$Region)
summary.aov(mod0)
summary.manova(mod0)
summary.manova(mod0, test="Wilks")
```
*The one way MANOVA suggests that Region is a statistically significant predictor for both Acceleration and Stamina*

#### Two-way Manova
```{r}
mod1 <- manova(as.matrix(PDeurlat[,c("Acceleration", "Stamina")]) ~ PDeurlat$Region + PDeurlat$InternationalReputation + PDeurlat$Region*PDeurlat$InternationalReputation) 

summary.aov(mod1)
summary.manova(mod1)
summary.manova(mod1, test="Wilks")
```
*The Two-Way MANOVA suggests that both Region and International Reputation (but not the interaction of the two) are statistically significant predictors for Acceleration and Stamina, respectively.*

#### Examining how skillMoves interacts with our response variables
```{r}
plot(PDeurlat$Wage, PDeurlat$Acceleration, main = 'Wage ~ Acceleration')
plot(PDeurlat$Wage, PDeurlat$Stamina, main = 'Wage ~ Stamina')
```

# GLM with skillMoves variable
```{r}
mod2 <- lm(as.matrix(PDeurlat[,c("Acceleration", "Stamina")]) ~ PDeurlat$Region + PDeurlat$InternationalReputation + PDeurlat$Wage +  PDeurlat$Region*PDeurlat$InternationalReputation) 

summary(mod2)
```

```{r}
source("http://www.reuningscherer.net/STAT660/R/CSQPlot.r.txt")
CSQPlot(mod1$residuals, label="MANOVA Residuals")
#maybe this should be changed to the GLM residuals instead of MANOVA? possibly just add GLM results?
CSQPlot(mod2$residuals, label="GLM Residuals")
```
[insert analysis here]

#Conclusions and Discussion

#Points for further analysis
